---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this short example, the aim is to predict Toyota Corolla prices by taking the fields such as age, kilometers, fuel type etc. into consideration. The tree will be pruned according to the cross-validation error. 

## Importing the Data and the Required Libraries

```{r intro1, warning=FALSE, echo=FALSE}

library(rpart)
library(rpart.plot)
library(caret)
library(tree)
library(caTools)
library(dplyr)
library(Metrics)

```


```{r intro2}

tc <- read.csv("ToyotaCorolla.csv")

seed <- 425

set.seed(seed)

# Partitioning the data set into training and test sets

split   <- sample.split(tc$Price, SplitRatio = 0.80)

tctrain <- subset(tc, split == TRUE)
tctest  <- subset(tc, split == FALSE)

nrow(tctrain)
nrow(tctest)

```

## Generating and Pruning the Tree

```{r tree}
tree <- rpart(Price~., data = tctrain)

prp(tree,
    type = 5,
    extra = 1,
    tweak = 1)

```


```{r cptable, warning=FALSE}

cpTable <- printcp(tree)

# Reporting the number of terminal nodes in the tree with the lowest cv-error, 
# which is equal to [the number of splits performed to create the tree] + 1

optIndex <- which.min(unname(tree$cptable[, "xerror"]))

cpTable[optIndex, 2] + 1

```

The generated tree has `r cpTable[optIndex, 2] + 1` terminal nodes.

```{r prune}

# Pruning the tree to the optimized cp value

optTree <- prune.rpart(tree = tree, cp = cpTable[optIndex, 1])

prp(optTree)

```

## Performing the Predictions and Reporting the Metrics

```{r pred}

# Making predictions in the test set

pred <- predict(optTree, newdata = tctest)

# Reporting the metrics

rmse(actual = tctest$Price, predicted = pred)
mae (actual = tctest$Price, predicted = pred)

```

